---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# centerline

<!-- badges: start -->
[![R-CMD-check](https://github.com/atsyplenkov/centerline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/atsyplenkov/centerline/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `centerline` R package simplifies the extraction of linear features from complex polygons, such as roads or rivers, by computing their centerlines (or median-axis) using Voronoi diagrams. It uses the super-fast [`geos`](https://paleolimbot.github.io/geos/index.html) and [`rmapshaper`](http://andyteucher.ca/rmapshaper/index.html) libraries in the background.

## Installation

You can install the development version of `centerline` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atsyplenkov/centerline")

# OR

# install.packages("pak")
pak::pak("atsyplenkov/centerline")
```

## Example

At the heart of this package is the `cent_skeleton` function, which efficiently computes the skeleton of closed 2D polygonal geometries. The function uses [`rmapshaper::ms_simplify()`](http://andyteucher.ca/rmapshaper/reference/ms_simplify.html) by default to keep the most important nodes and reduce noise from the beginning.

```{r example}
library(centerline)
library(terra)

polygon <- 
  terra::vect("inst/extdata/example.gpkg", layer = "polygon")

plot(polygon)

pol_skeleton <- 
  cent_skeleton(polygon, simplify = T) 

pol_skeleton |> 
  plot(col = "blue", add = T)

```

