---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%", 
  dpi = 1000
)
```

# centerline

<!-- badges: start -->
[![R-CMD-check](https://github.com/atsyplenkov/centerline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/atsyplenkov/centerline/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/centerline)](https://CRAN.R-project.org/package=centerline)
![GitHub R package version](https://img.shields.io/github/r-package/v/atsyplenkov/centerline?label=github)
![GitHub last commit](https://img.shields.io/github/last-commit/atsyplenkov/centerline)
<!-- badges: end -->

The `centerline` R package simplifies the extraction of linear features from complex polygons, such as roads or rivers, by computing their centerlines (or median-axis) using Voronoi diagrams. It uses the super-fast [`geos`](https://paleolimbot.github.io/geos/index.html) and [`rmapshaper`](http://andyteucher.ca/rmapshaper/index.html) libraries in the background.

## Installation

You can install the development version of `centerline` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atsyplenkov/centerline")

# OR

# install.packages("pak")
pak::pak("atsyplenkov/centerline")
```

## Examples for closed geometries

At the heart of this package is the `cnt_skeleton` function, which efficiently computes the skeleton of closed 2D polygonal geometries. The function uses [`rmapshaper::ms_simplify()`](http://andyteucher.ca/rmapshaper/reference/ms_simplify.html) by default to keep the most important nodes and reduce noise from the beginning. However, it has option to densify the amount of points using [`geos::geos_densify`](https://paleolimbot.github.io/geos/reference/geos_centroid.html), which can produce more smooth results.

```{r}
#| label: "Skeleton example"

library(centerline)
library(terra)

# Load Polygon Of Interest (POI)
polygon <- 
  terra::vect(
    system.file(
      "extdata/example.gpkg", package = "centerline"
    ), 
    layer = "polygon"
  )

# Find POI's skeleton
pol_skeleton <- 
  cnt_skeleton(polygon, keep = 1) 

# Simplified POI's skeleton
pol_skeleton_simplify <- 
  cnt_skeleton(polygon, keep = .1) 

# Densified POI's skeleton
pol_skeleton_densify <- 
  cnt_skeleton(polygon, keep = 1.5) 

```

```{r}
#| label: plot-1
#| code-fold: true
#| code-summary: "Plot's code"

# Plot
par(
  mar = c(0.5, 0.5, 0.2, 0.2),
  mfrow = c(1, 3),
  oma = c(0, 0, 0.2, 0.2)
)

# Raw
plot(pol_skeleton,
     col = "dodgerblue3")
plot(polygon,
     border = "grey20",
     main = "Original",
     add = T)

# Simplified
plot(pol_skeleton_simplify,
     col = "dodgerblue3")
plot(polygon,
     border = "grey20",
     main = "Simplified",
     add = T)

# Densified
plot(pol_skeleton_densify,
     col = "dodgerblue3")
plot(polygon,
     border = "grey20",
     main = "Densified",
     add = T)

```

However, the above-generated lines are not exactly a centerline of a polygon. One way to find the centerline of a closed polygon is to define both `start` and `end` points. For example, in the case of landslides, it could be the landslide initiation point and landslide terminus.

```{r}
#| label: "Connect points example"

# Load points data
points <- 
  terra::vect(
    system.file(
      "extdata/example.gpkg", package = "centerline"
    ),
    layer = "polygon_points"
  )

# Connect points
# For original skeleton
pol_path <-
  cnt_path(
    skeleton = pol_skeleton,
    start_point = terra::subset(points, points$type == "start"),
    end_point = terra::subset(points, points$type != "start")
  )

# For simplified skeleton
pol_path_simplify <-
  cnt_path(
    skeleton = pol_skeleton_simplify,
    start_point = terra::subset(points, points$type == "start"),
    end_point = terra::subset(points, points$type != "start")
  )

# For densified skeleton
pol_path_dens <-
  cnt_path(
    skeleton = pol_skeleton_densify,
    start_point = terra::subset(points, points$type == "start"),
    end_point = terra::subset(points, points$type != "start")
  )

```

```{r}
#| label: plot-2
#| code-fold: true
#| code-summary: "Plot's code"

# Plot
par(mfrow = c(1, 3)) 

# Original
plot(polygon, border = "grey20", 
     main = paste0("Original path (L = ",
                   round(terra::perim(pol_path[[1]]), 2), " m)"))
plot(pol_path[[1]], lwd = 3, add = T)
plot(points[1, ], col = "coral2",  add = T)
plot(points[2, ], col = "green4",  add = T)

# Simplified
plot(polygon, border = "grey20", 
     main = paste0("Simplified path (L = ",
                   round(terra::perim(pol_path_simplify[[1]]), 2), " m)"))
plot(pol_path_simplify[[1]], lwd = 3, add = T)
plot(points[1, ], col = "coral2",  add = T)
plot(points[2, ], col = "green4",  add = T)

# Densified
plot(polygon, border = "grey20", 
     main = paste0("Densified path (L = ",
                   round(terra::perim(pol_path_dens[[1]]), 2), " m)"))
plot(pol_path_dens[[1]], lwd = 3, add = T)
plot(points[1, ], col = "coral2",  add = T)
plot(points[2, ], col = "green4",  add = T)

```

## Roadmap
```
centerline ðŸ“¦
â”œâ”€â”€ Closed geometries (e.g., lakes)
â”‚   â”œâ”€â”€ When we do know starting and ending points (e.g., landslides) âœ…
â”‚   â”‚   â”œâ”€â”€ centerline::cnt_skeleton âœ…
â”‚   â”‚   â””â”€â”€ centerline::cnt_path âœ…
â”‚   â””â”€â”€ When we do NOT have points (e.g., lakes) ðŸ”²
â”‚       â”œâ”€â”€ centerline::cnt_skeleton âœ…
â”‚       â””â”€â”€ centerline::cnt_path_guess ðŸ”²
â”œâ”€â”€ Linear objects (e.g., roads or rivers)  ðŸ”²
â””â”€â”€ Collapse parallel lines to centerline ðŸ”²
```

## Alternatives
- **R**
	- [midlines](https://github.com/RichardPatterson/midlines) - A more hydrology-oriented library that provides a multi-step approach to generate a smooth centerline of complex curved polygons (like rivers).
	- [cmgo](https://github.com/AntoniusGolly/cmgo) - The main aim of the package is to propose a workflow to extract channel bank metrics, and as a part of that workflow, centerline extraction was implemented.
- ðŸ Python: 
	- [centerline](https://github.com/fitodic/centerline/tree/master) library
- ðŸ¦€ Rust: 
	- [centerline_rs](https://codeberg.org/eadf/centerline_rs) library
- **JS** Javascript:
  - [Centerline labeling blogpost](https://observablehq.com/@veltman/centerline-labeling)
