---
format: gfm
editor_options:
  chunk_output_type: console
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 1000
)

requireNamespace("geomtextpath", quietly = TRUE)
requireNamespace("smoothr", quietly = TRUE)
```

# centerline

<!-- badges: start -->
[![R-CMD-check](https://github.com/atsyplenkov/centerline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/atsyplenkov/centerline/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/centerline)](https://CRAN.R-project.org/package=centerline)
![GitHub R package version](https://img.shields.io/github/r-package/v/atsyplenkov/centerline)
![GitHub last commit](https://img.shields.io/github/last-commit/atsyplenkov/centerline)
[![Codecov test coverage](https://codecov.io/gh/atsyplenkov/centerline/graph/badge.svg)](https://app.codecov.io/gh/atsyplenkov/centerline)
<!-- badges: end -->

The `centerline` R package simplifies the extraction of linear features from complex polygons, such as roads or rivers, by computing their centerlines (or median-axis) using Voronoi diagrams. It uses the super-fast [`geos`](https://paleolimbot.github.io/geos/index.html) library in the background.

## Installation

You can install the development version of `centerline` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atsyplenkov/centerline")

# OR

# install.packages("pak")
pak::pak("atsyplenkov/centerline")
```

## Examples for closed geometries

At the heart of this package is the `cnt_skeleton` function, which efficiently computes the skeleton of closed 2D polygonal geometries. The function uses [`geos::geos_simplify`](https://paleolimbot.github.io/geos/reference/geos_centroid.html) by default to keep the most important nodes and reduce noise from the beginning. However, it has option to densify the amount of points using [`geos::geos_densify`](https://paleolimbot.github.io/geos/reference/geos_centroid.html), which can produce more smooth results. Otherwise, you can set the parameter `keep = 1` to work with the initial geometry.

```{r}
#| label: "Skeleton example"

library(centerline)
library(terra)

# Load Polygon Of Interest (POI)
polygon <-
  terra::vect(
    system.file(
      "extdata/example.gpkg", package = "centerline"
    ),
    layer = "polygon"
  )

# Find POI's skeleton
pol_skeleton <-
  cnt_skeleton(polygon, keep = 1)

# Simplified POI's skeleton
pol_skeleton_simplify <-
  cnt_skeleton(polygon, keep = .1)

# Densified POI's skeleton
pol_skeleton_densify <-
  cnt_skeleton(polygon, keep = 1.5)

```

```{r}
#| label: plot-1
#| code-fold: true
#| code-summary: "Plot's code"

# Plot
par(
  mar = c(0.5, 0.5, 0.2, 0.2),
  mfrow = c(1, 3),
  oma = c(0, 0, 0.2, 0.2)
)

# Raw
plot(pol_skeleton,
     col = "dodgerblue3")
plot(polygon,
     border = "grey20",
     main = "Original",
     add = T)

# Simplified
plot(pol_skeleton_simplify,
     col = "dodgerblue3")
plot(polygon,
     border = "grey20",
     main = "Simplified",
     add = T)

# Densified
plot(pol_skeleton_densify,
     col = "dodgerblue3")
plot(polygon,
     border = "grey20",
     main = "Densified",
     add = T)

```

However, the above-generated lines are not exactly a centerline of a polygon. One way to find the centerline of a closed polygon is to define both `start` and `end` points. For example, in the case of landslides, it could be the landslide initiation point and landslide terminus.

```{r}
#| label: "Connect points example"

# Load points data
points <-
  terra::vect(
    system.file(
      "extdata/example.gpkg", package = "centerline"
    ),
    layer = "polygon_points"
  )

# Connect points
# For original skeleton
pol_path <-
  cnt_path(
    skeleton = pol_skeleton,
    start_point = terra::subset(points, points$type == "start"),
    end_point = terra::subset(points, points$type != "start")
  )

# For simplified skeleton
pol_path_simplify <-
  cnt_path(
    skeleton = pol_skeleton_simplify,
    start_point = terra::subset(points, points$type == "start"),
    end_point = terra::subset(points, points$type != "start")
  )

# For densified skeleton
pol_path_dens <-
  cnt_path(
    skeleton = pol_skeleton_densify,
    start_point = terra::subset(points, points$type == "start"),
    end_point = terra::subset(points, points$type != "start")
  )

```

```{r}
#| label: plot-2
#| code-fold: true
#| code-summary: "Plot's code"

# Plot
par(mfrow = c(1, 3))

# Original
plot(polygon, border = "grey20",
     main = paste0("Original path (L = ",
                   round(terra::perim(pol_path[1, ]), 2), " m)"))
plot(pol_path[1, ], lwd = 3, add = T)
plot(points[1, ], col = "coral2",  add = T)
plot(points[2, ], col = "green4",  add = T)

# Simplified
plot(polygon, border = "grey20",
     main = paste0("Simplified path (L = ",
                   round(terra::perim(pol_path_simplify[1, ]), 2), " m)"))
plot(pol_path_simplify[1, ], lwd = 3, add = T)
plot(points[1, ], col = "coral2",  add = T)
plot(points[2, ], col = "green4",  add = T)

# Densified
plot(polygon, border = "grey20",
     main = paste0("Densified path (L = ",
                   round(terra::perim(pol_path_dens[1, ]), 2), " m)"))
plot(pol_path_dens[1, ], lwd = 3, add = T)
plot(points[1, ], col = "coral2",  add = T)
plot(points[2, ], col = "green4",  add = T)

```
And what if we don't know the starting and ending locations? What if we just want to place our label accurately in the middle of our polygon? In this case, one may find the `cnt_path_guess` function useful. It returns the line connecting the most distant points, i.e., the polygon's length. Such an approach is used in limnology for measuring [lake lengths](https://www.lakescientist.com/lake-shape/), for example.

```{r}
#| label: "Lake labelling"
#| warning: false
#| message: false
library(sf)

lake <-
  st_read(
    system.file("extdata/example.gpkg", package = "centerline"),
    layer = "lake",
    quiet = T
  ) |>
  st_cast("POLYGON")

lake_centerline <-
  cnt_path_guess(lake, keep = 1)

```

```{r}
#| label: plot-3
#| code-fold: true
#| code-summary: "Plot's code"
library(ggplot2)
library(geomtextpath)
library(smoothr)

lake_centerline_s <-
  lake_centerline |>
  st_simplify(dTolerance = 150)  |>
  smooth("chaikin")

cnt2 <-
  rbind(
    st_as_sf(lake_centerline_s),
    st_as_sf(lake_centerline_s)
  )

cnt2$lc <- c("black", NA_character_)
cnt2$ll <- c("", lake$name)

ggplot() +
  geom_sf(
    data = lake,
    fill = "#c8e8f1",
    color = NA
  ) +
  geom_textsf(
    data = cnt2,
    aes(
      linecolor = lc,
      label = ll
    ),
    color = "#458894",
    size = 8
  ) +
  scale_color_identity() +
  facet_wrap(~lc) +
  theme_void() +
  theme(strip.text = element_blank())
```


## Roadmap
```
centerline ðŸ“¦
â”œâ”€â”€ Closed geometries (e.g., lakes)
â”‚   â”œâ”€â”€ When we do know starting and ending points (e.g., landslides) âœ…
â”‚   â”‚   â”œâ”€â”€ centerline::cnt_skeleton âœ…
â”‚   â”‚   â””â”€â”€ centerline::cnt_path âœ…
â”‚   â””â”€â”€ When we do NOT have points (e.g., lakes) âœ…
â”‚       â”œâ”€â”€ centerline::cnt_skeleton âœ…
â”‚       â””â”€â”€ centerline::cnt_path_guess âœ…
â”œâ”€â”€ Linear objects (e.g., roads or rivers)  ðŸ”²
â””â”€â”€ Collapse parallel lines to centerline ðŸ”²
```

## Alternatives
- **R**
	- [midlines](https://github.com/RichardPatterson/midlines) - A more hydrology-oriented library that provides a multi-step approach to generate a smooth centerline of complex curved polygons (like rivers).
	- [cmgo](https://github.com/AntoniusGolly/cmgo) - The main aim of the package is to propose a workflow to extract channel bank metrics, and as a part of that workflow, centerline extraction was implemented.
- ðŸ Python:
	- [centerline](https://github.com/fitodic/centerline/tree/master) library
- ðŸ¦€ Rust:
	- [centerline_rs](https://codeberg.org/eadf/centerline_rs) library
- **JS** Javascript:
  - [Centerline labeling blogpost](https://observablehq.com/@veltman/centerline-labeling)
